A continuous urbanization index to assess phylogenetic adaptations in urban bird living

### 2018-05-14 ms.Rmd
A little work on results and discussion

Here, we improve upon current methods of assessing biological traits which predicts urbanization in birds by using a series of publicly available datasets available for Australian birds. Importantly, we developed and used a continuous urbanization index as a response variable in the analyses. We had two objectives. Our first objective was to investigate whether there was a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes]. Second, we developed models which were and were not phylogenetically constrained to assess species' traits that predict pre-adaptation to urban living, using continuous and categorical predictor variables. These results are especially critical in the future action plans of Australia's birds [@garnett2011action], given Australia is currently undergoing a period of rapid urban growth.

@@

### Phylogenetic signal
There was a distinct phylogenetic signal in our urbanization index measure (Fig. 3; Table 1).

**MAYBE DO SOME MEAN index by FAMILY PRESENTATIONS HERE????**

## Model results
### Non phylogenetic models
The most important predictors of urbanization in the non-phylogenetic global model (Fig. 4) were feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006). Urban birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and migratory status (t = 2.15, p = 0.032). Urban birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), feeding in pairs and flocks (t = -2.24, p =0.025), being granivorous (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), using grassland/shrubland habitats (t = -4.26, p < 0.001), being insectivorous (t = -3.96, p < 0.001), and feeding solitary (t = -2.90, p = 0.004).

@@

### Differences in non-phylogenetic model and phylogenetic model
The phylogenetic and non-phylogenetic models largely agreed with the relative rank of standardized predictor estimates, and the significance of individual predictors (Fig. 4). However, brain residual size was not significant in either of the non-phylogenetic models, but was significantly associated with urbanization in our phylogenetic models (Fig. 5). Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (larger parameter estimates) in the phylogenetic model compared with the non-phylogenetic model (Fig. 5). Lastly, although not significant, body size was minimally important for the non-phylogenetic model, but substantially more important for the phylogenetically controlled model (Fig. 4 & Fig. 5).

#### Phylogenetic uncertainty
**WE ASSESSED PHYLOGENETIC UNCERTAINY AND FOUND THAT....**

# DISCUSSION
Our continuous urbanization index showed a significant phylogenetic signal [@revell2008phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence, @ibanez2017global, @sol2017urbanisation, @la2018phylogenetic]. This contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of sample size considerations. Niche breadth, as measured by different types of generalism, was significantly associated with our urbanization index across both phylogenetic and non-phylogenetic analyses, confirming the notion that urban birds have broad environmental tolerance [@bonier2007urban]. Conversely, granivores and insectivores were consistently selected against in urban areas. Residual brain mass, and body size to an extent, was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade, the species with the largest residual brain mass and large body size are most suited to surviving in urban areas.

Our finding that ecological generalists are more readily observed in urban areas confirms that of other work [@mckinney2006urbanization, @bonier2007urban, @moller2009successful]. We also confirm work which has found a relationship between annual fecundity and urbanization in birds [@croci2008does, @moller2009successful]. While some studies have found a lack of evidence that relative brain size is critical in urban birds [@kark2007living, @evans2011makes], we confirm those which do support the hypothesis that brain size is heavily correlated with urban birds [@moller2009successful, @maklakov2011brains, @moller2015brain]. Furthermore, we do confirm that urbanization appears to select against insectivorous species [@chace2006urban, @evans2011makes]. While others have found that urbanization supports granivorous species [@chace2006urban, @kark2007living, @evans2011makes], we found the opposite trend, but this could be a result of relatively little supplementary feeding in Australian cities [@fuller2008garden]. We did not find any strong support for our hypotheses on different nesting types, but we did find that in general, species which tended to be more social were at an advantage to city living, confirming previous studies [@kark2007living].

Clearly, many traits have been used to identify a species' adaptability with urban-living (e.g., Appendix 1), but often times results differ among studies. Fully understanding the suite of specific species' traits which are associated with urban birds allows for a better understanding of how to best preserve and maintain avian biodiversity in urban areas [@evans2011makes]. As previously noted, a major shortcoming of previous studies has been the focus on discrete classifications of species based on their responses to urbanization. Our analysis used a total of 477 species (Appendix 4), more than any similar study, to the best of our knowledge. Furthermore, our models assessed the applicability of traits in relation to one another (i.e., we did not run separate models or statistical tests for each specific trait as previously done). The aforementioned differences in our study's results compared with others is likely due to increased sample size, as well as our use of a continuous measure of urbanization. Our results suggest that urbanization selects against specialist species [@sorace2009distribution], granivores, and insectivores. As such, future urban development aimed at increasing avian diversity should focus on maintaining a high diversity of habitat [@melles2003urban] both for foraging [@shochat2004linking], and nesting [@fernandez2001habitat]. A species-specific approach to assess particular resources necessary for specialist species may be necessary. Lastly, a focal point should be on increasing insect abundance and diversity in urban environments (CITATION), which will in turn increase the presence of insectivorours bird species in urban environments.

Global broad-scale empirical datasets are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global, @aronson2016hierarchical, @gutierrez2018citizen], in large part due to the growing use of citizen science data [@silvertown2009new] in combination with advancing remote sensing technology [@gorelick2017google]. We used a novel approach, over a broad spatial and temporal scale, to quantify an urbanization index in birds. We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role, @lepczyk2017global, @katewaleffect], but migratory status was considered as a predictor. We highlight that future studies, especially within migrant-dominated landscapes (i.e., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full, @marra2015call]. Further, future studies should look to expand this analysis outside of the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2018non]).

>>>>

Here, we improve upon current methods of assessing biological traits which predicts urbanization in birds by using a series of publicly available datasets. We developed and used a continuous urbanization index as a response variable in the analyses. We had two objectives: (1) investigate whether there was a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes], and (2) develop models which were and were not phylogenetically constrained to assess species' traits that predict pre-adaptation to urban living, using a suite of continuous and categorical predictor variables. These results are especially critical in the future action plans of Australia's birds [@garnett2011action], given Australia is currently undergoing a period of rapid urban growth.

@@

### Phylogenetic signal
There was a distinct phylogenetic signal in our urbanization index measure (Fig. 3; Table 1), as it was significantly different than random for 4 out of the 5 indices. Additionally, there was strong phylogenetic signal for potential predictor traits, especially body size and brain residual (Table 1). The result for body size-- $k=3.3$ --is among the strongest ever recorded measures of phylogenetic signal in the literature (see REFS) and for $k$ the statistic suggests much stronger signal than expected from Brownian motion. We found weaker signal for behavioral traits compared to morphological traits. For these traits, the estimates of phylogenetic signal were weaker than Brownian motion but significant in most cases (Table 1).

## Model results
### Non-phylogenetic models
The most important predictors of urbanization in the non-phylogenetic global model (Fig. 4) were feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006). Urban birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and migratory status (t = 2.15, p = 0.032). Urban birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), feeding in pairs and flocks (t = -2.24, p =0.025), being granivorous (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), using grassland/shrubland habitats (t = -4.26, p < 0.001), being insectivorous (t = -3.96, p < 0.001), and feeding solitary (t = -2.90, p = 0.004).

@@

### Differences in non-phylogenetic and phylogenetic models
The phylogenetic and non-phylogenetic models largely agreed with the relative rank of standardized predictor estimates, and the significance of individual predictors (Fig. 4). However, brain residual size was not significant in either of the non-phylogenetic models, but was significantly associated with urbanization in our phylogenetic models (Fig. 5). Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (larger parameter estimates) in the phylogenetic model compared with the non-phylogenetic model (Fig. 5). Lastly, although not significant, body size was minimally important for the non-phylogenetic model, but substantially more important for the phylogenetically controlled models (Fig. 4 & Fig. 5).

# DISCUSSION
Niche breadth, as measured by different types of generalism, was significantly associated with our urbanization index across both phylogenetic and non-phylogenetic analyses, confirming the notion that urban birds have broad environmental tolerance [@bonier2007urban]. Conversely, granivores and insectivores were consistently negatively associated with urbanization. Furthermore, we found a significant phylogenetic signal [@revell2008phylogenetic] of the urbanization index in our analysis, suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence, @ibanez2017global, @sol2017urbanisation, @la2018phylogenetic]. This contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of increased statistical power from our increased sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes]. Our results demonstrate that considering phylogeny in regards to this question is important.

The effect of treating species as non-independent by considering phylogeny in the context of the linear model was apparent. For example, there was an increase in parameter estimates of the most important parameters (e.g., clutch size and feeding habitat generalism; Figures 4 and 5), and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig. S2). Further, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass and largest body size are most suited to surviving in urban areas [@major2010museum]. Conversely, other predictor variables became less crucial in the phylogenetically controlled models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural). For those variables which became less important, it is likely that there may be particular clades that include a number of species that are associated with both that particular trait and urbanization [see @felsenstein1985confidence].

Our finding that ecological generalists are more readily observed in urban areas confirms that of other work [@mckinney2006urbanization, @bonier2007urban, @moller2009successful]. We also confirm work which has found a relationship between annual fecundity and urbanization in birds [@croci2008does, @moller2009successful]. While some studies have found a lack of evidence that relative brain size is critical in urban birds [@kark2007living, @evans2011makes], we confirm those which do support the hypothesis that brain size is heavily correlated with urban birds [@moller2009successful, @maklakov2011brains, @moller2015brain]. Furthermore, we do confirm that urbanization appears to select against insectivorous species [@chace2006urban, @evans2011makes]. While others have found that urbanization supports granivorous species [@chace2006urban, @kark2007living, @evans2011makes], we found the opposite trend, but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden]. We did not find any strong support for our hypotheses on different nesting types, but we did find that in general, species which tended to be more social were at an advantage to city living, confirming previous studies [@kark2007living].

Clearly, many traits have been used to identify a species' adaptability with urban-living (e.g., Appendix 1), but often times results differ among studies. Fully understanding the suite of specific species' traits which are associated with urban birds allows for a better understanding of how to best preserve and maintain avian biodiversity in urban areas [@evans2011makes]. As previously noted, a major shortcoming of previous studies has been the focus on discrete classifications of species based on their responses to urbanization. Our analysis used a total of 477 species (Appendix 4), more than any similar study, to the best of our knowledge. Furthermore, our models assessed the applicability of traits in relation to one another (i.e., we did not run separate models or statistical tests for each specific trait as previously done). The aforementioned differences in our results compared with others is likely due to increased sample size, as well as our use of a continuous measure of urbanization. Our results suggest that urbanization selects against specialist species [@sorace2009distribution], granivores, and insectivores. As such, future urban development aimed at increasing avian diversity should focus on maintaining a high diversity of habitat [@melles2003urban] both for foraging [@shochat2004linking], and nesting [@fernandez2001habitat]. A species-specific approach to assess particular resources necessary for specialist species may be necessary [@fernandez2004spatial]. Lastly, a focal point should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], which will in turn increase the presence of insectivorous bird species in urban environments.

Global broad-scale empirical datasets are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global, @aronson2016hierarchical, @gutierrez2018citizen], in large part due to the growing use of citizen science data [@silvertown2009new] in combination with advances in remote sensing technology [@gorelick2017google]. We used a novel approach, over a broad spatial and temporal scale, to quantify an urbanization index in birds. We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role, @lepczyk2017global, @katewaleffect], but migratory status was considered as a predictor. We highlight that future studies, especially within migrant-dominated landscapes (e.g., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full, @marra2015call]. Further, future studies should look to expand this analysis outside the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2018non]).
